---
title: "Fuel Efficiency and Engine Size in Classic Cars"
author: "Isi"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  ioslides_presentation:
    self_contained: true
    widescreen: true
---

```{r setup, include=FALSE}
.libPaths(c("r-lib", .libPaths()))
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

library(plotly)
library(dplyr)

cars <- tibble::rownames_to_column(mtcars, var = "model") %>%
  mutate(
    transmission = ifelse(am == 1, "Manual", "Automatic"),
    cylinders = factor(cyl)
  )
```

## Why this topic?

The `mtcars` dataset is small, recognizable, and perfect for a short web presentation.

- Fuel efficiency matters for both drivers and manufacturers.
- Engine size and horsepower often trade off against miles per gallon.
- An interactive chart makes those tradeoffs easy to explore.

## What the data includes

This presentation uses 32 car models from the built-in `mtcars` dataset.

- `mpg`: miles per gallon
- `disp`: engine displacement
- `hp`: horsepower
- `wt`: vehicle weight
- `am`: transmission type

## Interactive Plotly view

Hover over each point to inspect a specific car model.

```{r}
plot_ly(
  data = cars,
  x = ~disp,
  y = ~mpg,
  color = ~transmission,
  colors = c("#1b6ca8", "#d95f02"),
  size = ~hp,
  sizes = c(12, 36),
  text = ~paste(
    "<b>", model, "</b><br>",
    "Horsepower: ", hp, "<br>",
    "Weight: ", wt, " (1000 lbs)<br>",
    "Cylinders: ", cylinders
  ),
  hoverinfo = "text",
  type = "scatter",
  mode = "markers"
) %>%
  layout(
    title = "Fuel efficiency falls as engine displacement grows",
    xaxis = list(title = "Displacement"),
    yaxis = list(title = "Miles per gallon"),
    legend = list(title = list(text = "Transmission"))
  )
```

## Main takeaway

- Smaller engines tend to deliver better fuel economy.
- Manual cars often appear among the more efficient models.
- High-horsepower cars are concentrated in the low-MPG region.
